Red-Giant Hot Jupiters: Brilliant Radio Beacons?

### 2015-08-25 ms.tex
starting to finish the sensitivity section

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color}
%\usepackage[dvipdfmx]{color}
%\usepackage[dvipdfmx]{graphicx}

@@

In this section, we discuss the observability of the estimated radio emission. 
An obvious obstacle is the intrinsic radio emission from host red giant stars themselves; if they are larger than radio flux from the planets, there is no way to identify the latter. We will see in \S\ref{ss:RGradio} that radio flux from the planets will probably be larger in a certain parameter range. 
Then, we estimate the radio emission of Jupiter-twin around known M giants in \S\ref{ss:actualMgiants}. 
Finally, \S\ref{ss:timevariability} discusses the expected time variability of the radio emission as a mean to discern the signals from the background noise.

@@

\memoYF{discussion about detectability... Figures 1-3 of \citet{griesmeier2007b} show $10^{-3}$ Jy limit for 10-100 MHz (LWA, LOFAR) and $10^{-5}-10^{-4}$ Jy for 100-1000 MHz (LOFAR, SKA). }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & Time  & Resolution & Confusion limit\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy \\[0.25pc] 
\hline\\[-0.5pc]
%ASKAP & 700-1700~MHz & 37 & 342.3 \\ % table 3 of http://www.atnf.csiro.au/projects/askap/config.html
GMRT  & 235~MHz &  165.6 & 6860 & 13 \\ % http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
HERA  & 70~MHz  &  &  \\ % http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
LOFAR & ~45  &  833!! &  &  & \\
LOFAR & ~60  &  648!! &  &  & \\
LOFAR & 120  &  51.73 &  &  & \\  
LOFAR & 150  &  40.97 &  &  & \\  
LWA   & ~70  &  &  \\
SKA   & 200  &  &  \\
UTR-2 & & & \\
VLA P-band & 270-470  & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 2~$\mu$Jy RMS.  We consider only
telescopes that can reach this in less than XX hours without
being confusion limited.}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{deluxetable}{l|lll}
%\tablecolumns{4}
%\tablewidth{0pt}
%\tabletypesize{\scriptsize}
%\tablecaption{Detectability of Hot Jupiters with Current and Future Radio Telescopes}
%\tablehead{ Instrument & Band & Sensitivity & Time  \\ 
% &  & $\mu$Jy~hr$^{1/2}$ &  hr} 
%\startdata 
%GMRT & 220~MHz   &  & \\
%SKA  & 200~MHz)  &  &  \\
%VLA  & P-band (270-430~MHz)  &  & 
%\enddata
%\tablecomments{Time estimates to reach 10~$\mu$Jy RMS.  We consider only
%telescopes that can reach this in less than XX hours without
%being confusion limited.}
%\label{tab:sens}
%\end{deluxetable}

>>>>

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color}
\usepackage{url}
%\usepackage[dvipdfmx]{color}
%\usepackage[dvipdfmx]{graphicx}

@@

In this section, we discuss the observability of the estimated radio emission. 
An obvious obstacle is the intrinsic radio emission from host red giant stars themselves; if they are larger than radio flux from the planets, there is no way to identify the latter. We will see in \S\ref{ss:RGradio} that radio flux from the planets will probably be larger in a certain parameter range. 
Then, we estimate the radio emission of Jupiter-twin around known M giants in \S\ref{ss:actualMgiants}. 
In \S\ref{ss:detectability}, we estimate the sensitivities and limitations of several current and future
radio instruments at lower frequencies.
Finally, \S\ref{ss:timevariability} discusses the expected time variability of the radio emission as a mean to discern the signals from the background noise.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Detectability with Current and Near Future Low Frequency Radio Instrumentation}
\label{ss:detectability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\memoYF{discussion about detectability... Figures 1-3 of \citet{griesmeier2007b} show $10^{-3}$ Jy limit for 10-100 MHz (LWA, LOFAR) and $10^{-5}-10^{-4}$ Jy for 100-1000 MHz (LOFAR, SKA).}

We consider in this section several current and near future radio observatories that 
operate at frequencies $\nu \lesssim 300~\rm MHz$ and can reach continuum 
sensitivities less than 1~mJy in a few hours.  The results are summarized
in Table~\ref{tab:sens}.
The primary limitations to detectability are noise, characterized by the sky
and system equivalent blackbody temperatures $T_{\rm sky}$ and $T_{\rm sys}$,
resolution, and source confusion.  
To estimate the source confusion $\sigma_c$ (in mJy beam$^{-1}$) reported in Table~\ref{tab:sens}, 
we use equations provided by \cite{condon+ransom2016}, reproduced below:
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!,~(\theta < 10\arcsec)
\end{eqnarray}
The confusion limit is dominated by active galactic nuclei, which have low polarization 
fractions $<2.5\%$ \citep{Stil2014} and slowly vary with time \citep{Condon1974,Condon2012}.
Therefore both the circularly polarized nature of the emission from hot Jupiters 
and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate 
the source confusion limit.

The LOw-Frequency ARray (LOFAR) operates in the 10-240~MHz range, split in two arrays
using a low band and high band antenna design (LOFAR-LBA and LOFAR-HBA), 
and different pad configuration \cite{vanHaarlem2013}.  
LOFAR's two most sensitive bands are centered at 60~MHz for LOFAR-LBA and 150~MHz 
for LOFAR-HBA respectively.  Given the strong frequency dependence of
the effective aperture of a dipole antenna, and that the system noise temperature 
$T_{\rm sys}$ for LOFAR is not fully determined by the sky noise $T_{\rm sky}$, 
we use the LOFAR Image Noise calculator\footnote{
Heald; \url{http://www.astron.nl/~heald/test/sens.php}} for simplicity, which 
is based on SKA Memo 113\footnote{
 Nijboer, Pandey-Pommier, \& de Bruyn; SKA Memo 113; 
\url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, for
our fiducial sensitivity and imaging parameters in Table~\ref{tab:sens}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & Time  & Resolution & Confusion limit\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy bm$^{-1}$ \\[0.25pc] 
\hline\\[-0.5pc]
%ASKAP & 700-1700~MHz & 37 & 342.3 \\ % table 3 of http://www.atnf.csiro.au/projects/askap/config.html
GMRT  & 235~MHz &  165.6 & 6860 & 13 \\ % http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
HERA  & 70~MHz  &  &  \\ % http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
LOFAR &  30-80  & 700 &  & 8.3 & 120 \\
LOFAR & 110-200 &  35 &  & 3.3 & 2.6\\  
LWA   &  30-80  &  &  \\
SKA   & 200  &  &  \\
UTR-2 & & & \\
VLA P-band & 270-470  & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 2~$\mu$Jy RMS.}
\end{table*}
